---
title: "Make final datasets"
author: "Kyle Gontjes"
date: "`r Sys.Date()`"
output: html_document
---

# Updates: 

1. Remove absolute paths

2. Removed the 'major STs' category, cause that wasn't needed

```{r setup, include=FALSE}
require(knitr)
opts_knit$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA) 
opts_chunk$set(echo = T,warning = FALSE, message = FALSE,error=F,comment=NA)
opts_knit$set(root.dir='~/Desktop/gl_nfs/Project_INTEGRATE/Analysis/Resistance_dynamics/INTEGRATE-resistance-ms/')
```

# Environment

```{r environment, echo=T, message=T, include=T, results=T}
#Packages
packages <- c("tidyverse") 

#Load Packages
invisible(lapply(packages,library,character.only=T,quietly=T))

# Print environment
set.seed(45) 
Sys.info()
sessionInfo()
```

# Load merged dataset

```{r}
df_metadata_all <- readRDS("./data/dataset/CA_LTACH_CRE_genomes_merged_studies_metadata.RDS")
paste0("Total number of isolates: ",nrow(df_metadata_all))
paste0("Breakdown by study: ")
table(df_metadata_all$study)
```

# Clinical CRE isolates

1. Subset dataset to non-rectal isolates

```{r} 
df_metadata_all_clinical <- subset(df_metadata_all, rectal_logical==F)
saveRDS(df_metadata_all_clinical,"./data/dataset/CA_LTACH_clinical_CRE_genomes_merged_studies_metadata.RDS")
write_csv(df_metadata_all_clinical,"./data/dataset/CA_LTACH_clinical_CRE_genomes_merged_studies_metadata.csv")

# Summary
paste0("Total number of clinical CRE isolates: ",nrow(df_metadata_all_clinical))
paste0("Breakdown by study: ")
table(df_metadata_all_clinical$study)
```

# Clinical CRKP isolates

1. Subset dataset to Kpn isolates

2. Label the participant's first isolate

```{r} 
df_metadata_all_clinical_kpn <- subset(df_metadata_all_clinical,species == "Klebsiella pneumoniae")

# Get first clinical CRKP isolate using participant and date variable, but retain those with participant_id as unique entries 
df_metadata_all_clinical_kpn_first_isolate <- df_metadata_all_clinical_kpn %>% group_by(participant_id) %>% arrange(date) %>% slice(1) %>% ungroup() 

# First per quarter
df_metadata_all_clinical_kpn_first_isolate_per_Q <- df_metadata_all_clinical_kpn %>% group_by(participant_id,year_quarter) %>% arrange(date) %>% slice(1) %>% ungroup() 

# Create flag for in final dataset (i.e., first isolate or first quarter dataset)
df_metadata_all_clinical_kpn_first_isolate_per_Q$first_isolate <- df_metadata_all_clinical_kpn_first_isolate_per_Q$isolate_no %in% df_metadata_all_clinical_kpn_first_isolate$isolate_no
df_metadata_all_clinical_kpn$first_isolate <- df_metadata_all_clinical_kpn$isolate_no %in% df_metadata_all_clinical_kpn_first_isolate$isolate_no
df_metadata_all_clinical_kpn$first_isolate_per_quarter <- df_metadata_all_clinical_kpn$isolate_no %in% df_metadata_all_clinical_kpn_first_isolate_per_Q$isolate_no

# All clinical isolates
saveRDS(df_metadata_all_clinical_kpn,"./data/dataset/CA_LTACH_clinical_CRKP_genomes_merged_studies_metadata.RDS") 
write_csv(df_metadata_all_clinical_kpn,"./data/dataset/CA_LTACH_clinical_CRKP_genomes_merged_studies_metadata.csv") 

# Summary
paste0("Total number of clinical CRKP isolates: ",nrow(df_metadata_all_clinical_kpn))
paste0("Breakdown by study: ")
table(df_metadata_all_clinical_kpn$study)
```

# Add metadata to Klebsiella pneumoniae dataset

```{r}
create_Kpn_Species_ST_variable <- function(dataset){
  df_201415 <- dataset %>% subset(study == "2014-15")
  df_202123 <- dataset %>% subset(study == "2021-23")
  ST_over_1 <- table(dataset$Species_ST) %>% subset(.>1) %>% names
  Species_ST_recode <- ifelse(dataset$Species_ST %in% c(ST_over_1), dataset$Species_ST,"Singleton STs") %>% gsub("Klebsiella pneumoniae ","",.)  
  data.frame(Species_ST_recode)
}

df_metadata_all_clinical_kpn_first_isolate <- df_metadata_all_clinical_kpn_first_isolate %>% 
  cbind(create_Kpn_Species_ST_variable(df_metadata_all_clinical_kpn_first_isolate) )

df_metadata_all_clinical_kpn_first_isolate_per_Q <- df_metadata_all_clinical_kpn_first_isolate_per_Q %>% 
  cbind(create_Kpn_Species_ST_variable(df_metadata_all_clinical_kpn_first_isolate_per_Q) ) 
```

# First isolate per patient

```{r}
# First isolate  
saveRDS(df_metadata_all_clinical_kpn_first_isolate,"./data/dataset/CA_LTACH_first_clinical_CRKP_genomes_merged_studies_metadata.RDS") 
write_csv(df_metadata_all_clinical_kpn_first_isolate,"./data/dataset/CA_LTACH_first_clinical_CRKP_genomes_merged_studies_metadata.csv")
```

# First isolate per patient per calendar quarter

```{r}
# First isolate per quarter
saveRDS(df_metadata_all_clinical_kpn_first_isolate_per_Q,"./data/dataset/CA_LTACH_first_clinical_per_quarter_CRKP_genomes_merged_studies_metadata.RDS") 
write_csv(df_metadata_all_clinical_kpn_first_isolate_per_Q,"./data/dataset/CA_LTACH_first_clinical_per_quarter_CRKP_genomes_merged_studies_metadata.csv")
```

# Summary

```{r}
paste0("Total number of first-quarter clinical CRKP isolates: ", nrow(df_metadata_all_clinical_kpn_first_isolate_per_Q))

paste0("Breakdown by study and follow-up status: ")
table(df_metadata_all_clinical_kpn_first_isolate_per_Q$study,df_metadata_all_clinical_kpn_first_isolate_per_Q$first_isolate,dnn = c("Study","First isolate"))
```
 